Potential involvement of the Frank-Starling mechanism in reduced stroke volume variability observed in a murine model of heart failure with reduced ejection fraction

### 2023-03-29 Manuscript/manuscript.tex
update overleaf

We were interested in studying the variability of heart-cycle periods and stroke volumes. Therefore, we measured the above described Poincaré-plot indices of heart-cycle period variability for every experimental record, and averaged the results corresponding to the control and HFrEF groups. The outcomes are plotted in Fig. \ref{fig:fig02}C. Observe that no statistically significant difference exists between the indices corresponding to the control and HFrEF groups. This suggests that, under our experimental conditions, heart failure does not affect heart rate variability (HRV).

The same methodology was used to investigate the variability of stroke volumes. The results are shown in Fig. \ref{fig:fig02}D. The HFrEF group's SD1 and SD2 values are significantly lower than those of the control group. This, together with the absence of changes in HRV suggests that both types of variability are dissociated. Understanding why this happens is an intriguing question in itself.

>>>>

We were interested in studying the variability of heart-cycle periods and stroke volumes. To achieve this, we evaluated the Poincaré-plot indices of heart-cycle period variability (SD1 and SD2) for each experimental recording and averaged the corresponding results for the control and HFrEF groups. Representative Poincaré plots of cycle periods from a mouse in the control group and another in the HFrEF group are presented in Fig. \ref{fig:fig02}A, while the average values of SD1 and SD2 indices for both groups are depicted in Fig. \ref{fig:fig02}C. Notably, we observed no statistically significant difference between the indices associated with the control and HFrEF groups. This suggests that, in our experimental setup, heart failure does not impact heart rate variability (HRV).

The methodology used to investigate stroke volume variability was identical to that used for heart-cycle period variability. Fig. \ref{fig:fig02}B displays representative Poincaré plots of stroke volumes for a mouse in the control group and another in the HFrEF group. Fig. \ref{fig:fig02}D shows the mean values of SD1 and SD2 for both groups. The HFrEF group exhibited significantly lower SD1 and SD2 values than the control group. This finding, combined with the lack of changes in HRV, indicates that the two types of variability are dissociated. The reason behind this dissociation is an intriguing question that deserves further investigation.
